Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations

### 2022-02-01 analysis/paper/paper.Rmd
adund and stomach results edits

spring_abund = all_est %>%
  filter(model == "Schnabel") %>%
  # filter(model == "Chapman") %>%
  select(-range, -prop_range) %>%
  inner_join(total_cpue,
             by = c("event_name", "species")) %>%
  select(species,
         model,
         model_event = event_name,
         N_F = N,
         Lci_F = Lci,
         Uci_F = Uci,
         cpue_F = catch_per_hr) %>%
  group_by(species, model) %>%
  summarise(across(c(N_F:cpue_F),
                   mean),
            .groups = "drop") %>%
    full_join(total_cpue %>%
              filter(str_detect(event_name, "Spring")) %>%
              select(species,
                     cpue_S = catch_per_hr),
            by = "species") %>%
  # calculate spring abundance estimate
  mutate(N_S = N_F * (cpue_S / cpue_F),
         Lci_S = Lci_F * (cpue_S / cpue_F),
         Uci_S = Uci_F * (cpue_S / cpue_F))

@@

## Potential Impacts to Chinook Salmon Populations

@@

Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish. The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)). During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)).For both estimates, the multiple census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (***figure reference***). Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimator was determined to provide the best estimates of abundance. Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`).<!--Need to add the delayed-mixing Schnabel estimate results--> For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean of the fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough. Those estimates translate to linear densities of ***value*** and ***value*** Northern Pikeminnow per RKM.

<!-- need to adjust the spring estimate -->

The lengths of Northern Pikeminnow captured ranged from ***value*** to ***value*** mm with an average length of ***value*** mm (\@ref(fig:length-freq-p)). The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.

@@

We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective. Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs. Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)). The average wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r round(min(stmch_df$stomach_contents_weight), 2)`-`r round(max(stmch_df$stomach_contents_weight), 2)` g). Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.

@@

```{r adult-imp-tab, echo =TRUE}

impact_tbl = impact_df %>%
  select(season,
         perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c("N", "Lci", "Uci"),sum),
            .groups = "drop") %>%
  mutate(Stage = "Juveniles") %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  filter(perc_chnk %in% c(30, 50, 65))%>%

>>>>

spring_abund = all_est %>%
  filter(model == "Schnabel - Delayed Mixing") %>%
  # filter(model == "Chapman") %>%
  select(-range, -prop_range) %>%
  inner_join(total_cpue,
             by = c("event_name", "species")) %>%
  select(species,
         model,
         model_event = event_name,
         N_F = N,
         Lci_F = Lci,
         Uci_F = Uci,
         cpue_F = catch_per_hr) %>%
  group_by(species, model) %>%
  summarise(across(c(N_F:cpue_F),
                   mean),
            .groups = "drop") %>%
    full_join(total_cpue %>%
              filter(str_detect(event_name, "Spring")) %>%
              select(species,
                     cpue_S = catch_per_hr),
            by = "species") %>%
  # calculate spring abundance estimate
  mutate(N_S = N_F * (cpue_S / cpue_F),
         Lci_S = Lci_F * (cpue_S / cpue_F),
         Uci_S = Uci_F * (cpue_S / cpue_F))

@@

## Impacts to Chinook Salmon Populations

@@

Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish. The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021. During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)). For both surveys, the multiple census estimates were generally larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)). Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, the Schnabel estimates were determined to provide the best estimates of abundance. Additionally, due to concerns about marked and released fish not immediately mixing back into preferred angling locales and/or being "hook shy", we used the conservative delayed-mixing Schnabel estimates. Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`). For spring 2021, we estimated abundance by multiplying the mean fall abundance by the ratio between the spring CPUE and mean fall CPUEs (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), resulting in an estimate of `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough. Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm.

The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm with an average TL of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)). The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.

@@

We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective. Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs. Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)). The mean wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; sd = `r round(sd(stmch_df$stomach_contents_weight), 2)` g). Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.

@@

```{r adult-imp-tab, echo =TRUE}
impact_tbl = impact_df %>%
  select(season,
         perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c("N", "Lci", "Uci"),sum),
            .groups = "drop") %>%
  mutate(Stage = "Juveniles") %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  filter(perc_chnk %in% c(30, 50, 65))%>%
